---
title: "Trial 8 Night Max Is Zero"
author: "Abby Cannon"
date: "2024-10-18"
output: html_document
---

Let's start by bringing in our data
```{r}
datos <- read.csv("CorrectedTrial8NoBG.csv")
```

Now let's get just our night data
```{r}
class(datos$Time.of.Day) #It's character
#And let's check to make sure there are only 2 categories
unique(datos$Time.of.Day) #Only 2 types, which is perfect
noche <- datos[ which(datos$Time.of.Day == "night"), ]
dia <- datos[ which(datos$Time.of.Day == "day"), ]

#And now we check that dia didn't end up with any hours it shouldn't
unique(dia$Hour) #Doesn't look like it did
```

Let's get the nighttime maxima for each chamber
```{r}
library(dplyr) #The package we need to find max by chamber

#Getting our new data
night.max <- noche %>%
  group_by(Chamber) %>%
  summarise(max = max(MO2.no.Background.Correction, na.rm=TRUE))
```
So now we have no positive nighttime values, which is actually great.

Now let's rename a column for easier combining
```{r}
colnames(night.max)[2] <- "Max.Dark.MO2"
```

And come to think of it we probably don't have to be worried when the maximum nighttime value is less than zero. So we won't change those ones even though this means we somewhat underestimate production and overestimate consumption
```{r}
night.max$Max.Dark.MO2[night.max$Max.Dark.MO2<0] <- 0
```

Now let's put it back together with the data we already calculated
```{r}
calc <- merge(datos, night.max, by = "Chamber")
```

Now let's calculate MO2 as offset from Max.Dark.MO2
```{r}
calc$MO2.from.Dark.Max <- calc$MO2.no.Background.Correction - calc$Max.Dark.MO2
```

And now we apparently have more category names than we should, but that's also fixable.
```{r}
unique(calc$Treatment) #Alright. Looks like we need to get rid of some excess white
#space
library(stringr) #The package we need 
calc$Treatment <- str_trim(calc$Treatment, side = c("both"))
unique(calc$Treatment) #Better
```

So I think I should plot it to see what it looks like
```{r}
#We'll need ggplot for graphing and dplyr for knitting
library(ggplot2)
library(dplyr)

#Our plot
dibujo <- ggplot(calc, aes(x= Hour, y= MO2.from.Dark.Max, 
                 color= as.factor(Treatment))) +
  geom_tile(aes(x=Hour, y= 0.015, color = NULL, fill= Time.of.Day), width = 1, height = Inf, alpha = 0.05) +
  #makes tiles that are 1 unit wide and with their fill controlled by time of day. alpha of 0.05 makes them
  #somewhat translucent. Color=Null means no outlines only fill. Inf height means they fill the whole graph.
  scale_fill_manual("Time of Day", values=c("white", "gray90"), 
                    guide = guide_legend(override.aes = list(alpha = 1 )))+
  geom_hline(yintercept = 0, color = "black", linetype = "dashed")+
  #Lets us control what colors the fill are. Also lets us make the legend more intense than the treatment
  #bars
  geom_point(size = 0.5)+
  #makes the points smaller, which is good, because there are a lot of them
  scale_color_manual("Treatment", 
                     values = c("black", "#000080", "#FF0000", "#00CD00"))+
  #Lets us control the point and line color, which is nice
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 11), se = FALSE)+
  #Gives us smooth lines with some accounting for cycles
  #scale_y_continuous(limits = c(-0.1, 0.15))+
  #makes the scale consistent with other graphs
    ylab("Net Oxygen Production mg/hr/g")+
  xlab("Hour of Experiment")+
  ggtitle("Net Oxygen Production Over Time by Treatment")+
  theme_bw()
```

And let's print that graph
```{r}
dibujo
```

saves our graph, which is useful for the future
```{r}
ggsave(dibujo, 
       filename = "Trial8CorrectedRespirometry.png",
       device = "png")
```