---
title: "GNET Phase 2"
author: "Abby Cannon"
date: "2024-11-19"
output: html_document
---

```{r}
datos <- read.csv("GNET_Phase2.csv") #bring in our data
```

So we'll have to put in the trial combinations
```{r}
genes <- read.csv("TrialandGenotypes.csv") #another csv
```

And now we put them together
```{r}
datos2 <- merge(genes, datos, by = "Genotype")
```

And now let's get rid of the random x
```{r}
datos3 <- subset(datos2, select=-c(X))
```

Now let's find out what kind of data we have
```{r}
class(datos3$Trial) #integer
class(datos3$Day) #integer
class(datos3$Colletion.Time) #character and that's a problem. It also has a 
#typo in the column name so I may need to rename it for the sake of my sanity 
#and faster programming
```

So let's start with the renaming
```{r}
names(datos3)[names(datos3) == 'Colletion.Time'] <- 'Collection.time'
```

Now we make it the kind of data we need to be
So we try to fix collection time
```{r}
hora <- format(as.POSIXlt(datos3$Collection.time, format="%H:%M:%OS"),"%H:%M:%OS2") #So that worked even if it gave us more 
#precision than we strictly need
datos3$Collection.time <- hora
```

Now let's work on our turbidity (currently called treatment). We can start by renaming for the sake of our sanity
```{r}
names(datos3)[names(datos3) == 'Treatment'] <- 'Target.NTU'
```

And now let's check what kind of data we have for it and temperature
```{r}
class(datos3$Target.NTU) #Integer. Works.
class(datos3$Temperature) #Also integer. Also works.
```

Now we need to make an ID column to account for random effects of chamber/treatment combinations
```{r}
datos3$ID <- paste(datos3$Trial, datos3$Chamber, sep = "-")
datos3$ID <- as.factor(datos3$ID) #Need to make it a factor for subsequent
#analysis
```

Alright. I think now we can do some Bayesian analysis.
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(GNET ~ Day * Collection.time * 
                            Target.NTU * Temperature + ID,
                          data = datos3, whichRandom = "ID", neverExclude="^ID$")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm
```
What happens when we plot it?
```{r}
plot(bayes_rm)
```

So ID is [167] ID  : 0.1674266    ±0% Target NTU + ID is the most likely model that includes a variable and it's 0.03566071, which is less likely so we're done here.

