---
title: "GNET Phase 1"
author: "Abby Cannon"
date: "2024-11-18"
output: html_document
---
```{r}
datos <- read.csv("GNET_Phase1.csv") #bring in our data
```

So we'll have to put in the trial combinations
```{r}
genes <- read.csv("TrialandGenotypes.csv") #another csv
```

And now we put them together
```{r}
datos2 <- merge(genes, datos, by = "Genotype")
```

And now let's get rid of the random x
```{r}
datos3 <- subset(datos2, select=-c(X))
```

Now let's find out what kind of data we have
```{r}
class(datos3$Trial) #integer
class(datos3$Day) #integer
class(datos3$Collection.time) #character and that's a problem
```

So we try to fix collection time
```{r}
hora <- format(as.POSIXlt(datos3$Collection.time, format="%H:%M:%OS"),"%H:%M:%OS2") #So that worked even if it gave us more 
#precision than we strictly need
datos3$Collection.time <- hora
```

And now let's see what treatment is
```{r}
class(datos3$Treatment) #Integer should be just fine
```

Now we need to make an ID column to account for random effects of chamber/treatment combinations
```{r}
datos3$ID <- paste(datos3$Trial, datos3$Chamber, sep = "-")
datos3$ID <- as.factor(datos3$ID) #Need to make it a factor for subsequent
#analysis
```

Alright. I think now we can do some Bayesian analysis.
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```


```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(GNET ~ Day * Treatment * Collection.time + ID,
                          data = datos3, whichRandom = "ID", neverExclude="^ID$")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm
```
What happens when we plot it?
```{r}
plot(bayes_rm)
```
So nobody is more likely than ID and I guess that's where we can leave it.
