---
title: "Chla Together"
author: "Abby"
date: "2024-11-29"
output: html_document
---
Alright. Let's bring in our data
```{r}
bio <- read.csv("All_Bio_Data_Together.csv")
```

Now let's make a column for genotype, because that's important. We can do this by extracting everything in front of the dash.
```{r}
bio$Genotype <- sub("(^[^-]+)-.*", "\\1", bio$Coral.ID)
```

Alright. So I think there's a way to do a lot of grouped subtraction very quickly, which is helpful.
```{r}
library(dplyr) #The package we need to mutate

#Makes us a data frame where everything has the sacrifice subtracted from it by genotype
bio.cambios <- bio %>%
  group_by(Genotype) %>%
  mutate(change.symbiont = Symbiont.Counts..symbiont.cm2. - Symbiont.Counts..symbiont.cm2.[Treatment == "None"], change.chla = Chl.a - Chl.a[Treatment == "None"], change.Chl.C.2 = Chl.C.2 - 
           Chl.C.2[Treatment == "None"], change.Chl.a.C.2 = Chl.a.C.2 - 
          Chl.a.C.2[Treatment == "None"], change.Protein.Concentration = 
           Protein.Concentration..mg.cm.2.mL.. - Protein.Concentration..mg.cm.2.mL..[Treatment == "None"], 
         change.Chl.per.symbiont = Chl.a.c.2.per.symbiont..x1000. - Chl.a.c.2.per.symbiont..x1000.[Treatment == "None"])
```

Now we need to get rid of the trials we didn't end up using
```{r}
class(bio.cambios$Experiment) #Numeric, we can make that work
#New data frame without the experiment numbers we don't like
cambios <- bio.cambios[ which(bio.cambios$Experiment != 1.1 & 
                                bio.cambios$Experiment != 2.5), ]
#This also dropped all the NAs, but that may be okay
```

Now let's get our experiment numbers since I think that might be importance for difference from sacrifice
```{r}
#So first we need to account for 2.10 getting changed to 2.1, but that's an easy fix
cambios$Experiment <- as.character(cambios$Experiment)
cambios$Experiment <- gsub("2.1", "2.10", cambios$Experiment)

#Now let's get a column for just trial numbers
cambios$Trial.Sequence <- as.numeric(sub(".*\\.", "", cambios$Experiment))
```

And we may also want to get phase. I swear there might be something here
```{r}
cambios$Phase <- substr(cambios$Experiment, 1, 1)
```

So now let's work out what treatment is so we can color code our plots
```{r}
class(cambios$Treatment) #It's character, which I think should work well
unique(cambios$Treatment)
```

And now we need to get our turbidities and temperature
```{r}
#So turbidities are easy, temperatures may be a bit cheekier
cambios$Turb <- substr(cambios$Treatment, 1, 2)
cambios$Turb <- as.numeric(cambios$Turb)

#Let's see if we can reverse substring and then replace
library(stringr) #bringing in a package we need to substring backwards
cambios$Temp <- str_sub(cambios$Treatment,-3,-1) #Getting the last 3 characters or less when
#there are less
cambios$Temp <- gsub("15", "27", cambios$Temp)
cambios$Temp <- gsub("29", "27", cambios$Temp)
cambios$Temp <- gsub("50", "27", cambios$Temp)
cambios$Temp <- gsub("C", "", cambios$Temp)
cambios$Temp <- gsub("0", "27", cambios$Temp)
cambios$Temp <- gsub("327", "30", cambios$Temp) #It's only stupid if it doesn't work
cambios$Temp <- as.numeric(cambios$Temp)
```

Alright. Now I guess we do some Bayesian analysis of trials together
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

Now I just want column names for the sake of my sanity
```{r}
colnames(cambios)
```

And we need to make genotype random
```{r}
cambios$Genotype <- as.factor(cambios$Genotype)
cambios$Phase <- as.factor(cambios$Phase)
```

```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(Chl.a ~ Turb * 
                            Temp * Phase + Genotype, 
                          data = cambios, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm
```
So turbidity apparently matters and temperature doesn't.
127.8529/91.84336 = 1.39

And we'll need to tidy up the categories for a prettier graph.
```{r}

library(dplyr)
cambios <- cambios %>% rename(Treatment.bad = Treatment)

step.1 <- paste(cambios$Turb, cambios$Temp, sep = "NTU, ")
step.2 <- paste(step.1, "ºC", sep = "")
cambios$Treatment <- step.2 #Makes it the column we want to use for the graph
sort(unique(cambios$Treatment))
class(cambios$Treatment)
```

So now let's graph it and see what we get
```{r}
#And I guess we need to bring these in if we want the thing to work
source("R_rainclouds.R") #With the sources we need to copy them into the same project for unfortunate reasons
library(cowplot)
library(readr)
library(ggplot2)
library(dplyr)
library(ggrepel)


#And then we make our graph
p2 <- ggplot(cambios, aes(x = Turb, y = Chl.a, label = Genotype))+
  geom_flat_violin(data= cambios, aes(x= Turb, 
                                          y = Chl.a, 
                                        fill = Treatment), 
                   position = position_nudge(x = 0, y = 0), adjust = 1.5, trim = TRUE, alpha = .5, colour = NA)+
  geom_text_repel(aes(x= Turb, y = Chl.a, 
                                    colour = Treatment), 
                  nudge_x = -0.3, show.legend = FALSE)+
  scale_color_manual(values = c("black", "#FF0000", "#1E90FF", "#000080", 
                                "#00CD00", "#881B1B"))+
  scale_fill_manual(values = c("black", "#FF0000", "#1E90FF", "#000080", 
                                "#00CD00", "#881B1B"))+
  #theme_cowplot()+
  theme(panel.background = element_rect(fill = "white", colour = "black"))+
  #ylim(0.62, 0.725)+
  scale_x_continuous(breaks = c(0, 15, 29, 50), labels = c(0, 15, 29, 50), 
                     limits = c(-2, 57))+
  labs(fill = "Treatment")+
  xlab("Turbidity (NTU)")+
  ylab("Chlorophyll a μg/ml/cm^2")+
  ggtitle("Chlorophyll a by Turbidity")

p2
```
Alright. So maybe there is a threshold effect

```{r}
ggsave(p2, 
       filename = "ChlaCombined.png",
       device = "png")
```

What about when we look at offset
```{r}
#So now we make our Bayesian Models
bayes_rm.change.chla <- generalTestBF(change.chla ~ Turb * 
                            Temp*Phase + Genotype, 
                          data = cambios, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm.change.chla
```

```{r}
plot(bayes_rm.change.chla)
```
So more likely than just genotype, but not always

So now let's graph it and see what we get
```{r}
#And I guess we need to bring these in if we want the thing to work
source("R_rainclouds.R") #With the sources we need to copy them into the same project for unfortunate reasons
library(cowplot)
library(readr)
library(ggplot2)
library(dplyr)
library(ggrepel)


#And then we make our graph
p3 <- ggplot(cambios, aes(x = Turb, y = change.chla, label = Genotype))+
  geom_flat_violin(data= cambios, aes(x= Turb, 
                                          y = change.chla, 
                                        fill = Treatment), 
                   position = position_nudge(x = 0, y = 0), adjust = 1.5, trim = TRUE, alpha = .5, colour = NA)+
  geom_text_repel(aes(x= Turb, y = change.chla, 
                                    colour = Treatment), 
                  nudge_x = -0.3, show.legend = FALSE)+
  scale_color_manual(values = c("black", "black", "#FF0000", "#1E90FF", 
                                "#000080", "#000080", "#00CD00", "#881B1B"))+
  scale_fill_manual(values = c("black", "black", "#FF0000", "#1E90FF", 
                                "#000080", "#000080", "#00CD00", "#881B1B"))+
  #theme_cowplot()+
  theme(panel.background = element_rect(fill = "white", colour = "black"))+
  #ylim(0.62, 0.725)+
  scale_x_continuous(breaks = c(0, 15, 29, 50), labels = c(0, 15, 29, 50), 
                     limits = c(-2, 57))+
  labs(fill = "Treatment")+
  xlab("Turbidity (NTU)")+
  ylab("Change in Chlorophyll a")+
  ggtitle("Change Chlorophyll a by Turbidity")

p3
```


