---
title: "Protein Together"
author: "Abby Cannon"
date: "2024-11-30"
output: html_document
---

Alright. Let's bring in our data
```{r}
bio <- read.csv("All_Bio_Data_Together.csv")
```

Now let's make a column for genotype, because that's important. We can do this by extracting everything in front of the dash.
```{r}
bio$Genotype <- sub("(^[^-]+)-.*", "\\1", bio$Coral.ID)
```

Alright. So I think there's a way to do a lot of grouped subtraction very quickly, which is helpful.
```{r}
library(dplyr) #The package we need to mutate

#Makes us a data frame where everything has the sacrifice subtracted from it by genotype
bio.cambios <- bio %>%
  group_by(Genotype) %>%
  mutate(change.symbiont = Symbiont.Counts..symbiont.cm2. - Symbiont.Counts..symbiont.cm2.[Treatment == "None"], change.chla = Chl.a - Chl.a[Treatment == "None"], change.Chl.C.2 = Chl.C.2 - 
           Chl.C.2[Treatment == "None"], change.Chl.a.C.2 = Chl.a.C.2 - 
          Chl.a.C.2[Treatment == "None"], change.Protein.Concentration = 
           Protein.Concentration..mg.cm.2.mL.. - Protein.Concentration..mg.cm.2.mL..[Treatment == "None"], 
         change.Chl.per.symbiont = Chl.a.c.2.per.symbiont..x1000. - Chl.a.c.2.per.symbiont..x1000.[Treatment == "None"])
```

Now we need to get rid of the trials we didn't end up using
```{r}
class(bio.cambios$Experiment) #Numeric, we can make that work
#New data frame without the experiment numbers we don't like
cambios <- bio.cambios[ which(bio.cambios$Experiment != 1.1 & 
                                bio.cambios$Experiment != 2.5), ]
#This also dropped all the NAs, but that may be okay
```

Now let's get our experiment numbers since I think that might be importance for difference from sacrifice
```{r}
#So first we need to account for 2.10 getting changed to 2.1, but that's an easy fix
cambios$Experiment <- as.character(cambios$Experiment)
cambios$Experiment <- gsub("2.1", "2.10", cambios$Experiment)

#Now let's get a column for just trial numbers
cambios$Trial.Sequence <- as.numeric(sub(".*\\.", "", cambios$Experiment))
```

And we may also want to get phase. I swear there might be something here
```{r}
cambios$Phase <- substr(cambios$Experiment, 1, 1)
```

So now let's work out what treatment is so we can color code our plots
```{r}
class(cambios$Treatment) #It's character, which I think should work well
unique(cambios$Treatment)
```

And now we need to get our turbidities and temperature
```{r}
#So turbidities are easy, temperatures may be a bit cheekier
cambios$Turb <- substr(cambios$Treatment, 1, 2)
cambios$Turb <- as.numeric(cambios$Turb)

#Let's see if we can reverse substring and then replace
library(stringr) #bringing in a package we need to substring backwards
cambios$Temp <- str_sub(cambios$Treatment,-3,-1) #Getting the last 3 characters or less when
#there are less
cambios$Temp <- gsub("15", "27", cambios$Temp)
cambios$Temp <- gsub("29", "27", cambios$Temp)
cambios$Temp <- gsub("50", "27", cambios$Temp)
cambios$Temp <- gsub("C", "", cambios$Temp)
cambios$Temp <- gsub("0", "27", cambios$Temp)
cambios$Temp <- gsub("327", "30", cambios$Temp) #It's only stupid if it doesn't work
cambios$Temp <- as.numeric(cambios$Temp)
```

Alright. Now I guess we do some Bayesian analysis of trials together
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

Now I just want column names for the sake of my sanity
```{r}
colnames(cambios)
```

And we need to make genotype random
```{r}
cambios$Genotype <- as.factor(cambios$Genotype)
cambios$Phase <- as.factor(cambios$Phase)
```

```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(Protein.Concentration..mg.cm.2.mL.. ~ Turb * 
                            Temp * Phase + Genotype, 
                          data = cambios, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```
Prints out our model and tells us what's good.
```{r}
bayes_rm
```
All less likely than just genotype.

What about when we look at offset
```{r}
#So now we make our Bayesian Models
bayes_rm.change.protein<- generalTestBF(change.Protein.Concentration ~ Turb * 
                            Temp*Phase + Genotype, 
                          data = cambios, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```
Prints out our model and tells us what's good.
```{r}
bayes_rm.change.protein
```

This is a little more confusing so let's plot it.

```{r}
plot(bayes_rm.change.protein)
```
Also less likely than genotype.



