---
title: "ETRm After Experiment Phase 1"
author: "Abby Cannon"
date: "2024-11-01"
output: html_document
---

```{r}
#Let's bring in our data
datos <- read.csv("OFAV_PAM_P1_11Jun2024.csv")

#Looks good, but we need to get rid of the sacrifices
class(datos$Treatment) #Character
unique(datos$Treatment) #We need to get rid of "x" 

#Making a new dataframe without the sacrifice treatments
datos.no.sac <- datos[ which(datos$Treatment != "x"), ]

#Now we need to make our days more quantitative
dias <- read.csv("RecoveryPhaseDays.csv")

datos.days <- merge(datos.no.sac, dias, by = "Recovery.Phase")
```

Then we select for only the After data
```{r}
after.datos <- datos.days[ which(datos.days$Days.Post.Experiment == 0), ]
```

We might be able to leave treatment as is, depending on how it's classified
```{r}
class(after.datos$Treatment) #It's character. That's a problem.
Treatment <- as.numeric(after.datos$Treatment) #That's how we fix it
after.datos$Treatment <- Treatment
```

And now that we know that the replicates are of the same coral fragment, we should put them together to avoid pseudoreplication
```{r}
fluoro <- aggregate(Fv.Fm ~ Coral.ID + Treatment + Genotype, 
                    data = after.datos, mean)
```

Now it's time to start playing with data
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

And we should probably also consider Genotype as a potential factor to avoid pseudoreplication
```{r}
#And it needs to be a factor
fluoro$Genotype <- as.factor(fluoro$Genotype)
```


```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(Fv.Fm ~ Treatment + Genotype, 
                          data = fluoro, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm
```

```{r}
plot(bayes_rm)
```

Okay. Very similar

```{r}
#Gets every model diveded by evey other model
bfmat <- bayes_rm/bayes_rm
look.at.this <- bfmat[,2] #Prints out the 19th column, which should show everything divided by ID

look.at.this
```
I really don't find that compelling. Maybe we'll see what it looks like pre too.
It is more likely than pre, so I guess we can see what we got.
It is technically more likely than pre, but way too close to equal.

```{r}
#And I guess we need to bring these in if we want the thing to work
source("R_rainclouds.R") #With the sources we need to copy them into the same project for unfortunate reasons
library(cowplot)
library(readr)
library(ggplot2)
library(dplyr)
library(ggrepel)


#And then we make our graph
p2 <- ggplot(fluoro, aes(x = Treatment, y = ETRm, label = Genotype))+
  geom_flat_violin(data= fluoro, aes(x=Treatment, 
                                          y = ETRm, 
                                        fill = factor(Treatment)), 
                   position = position_nudge(x = 0, y = 0), adjust = 1.5, trim = TRUE, alpha = .5, colour = NA)+
  geom_text_repel(aes(x=Treatment, y = ETRm, 
                                    colour = factor(Treatment)), 
                  nudge_x = -0.5, show.legend = FALSE)+
  scale_color_manual(values = c("black", "#1E90FF", "#000080", "#881B1B"))+
  scale_fill_manual(values = c("black", "#1E90FF", "#000080", "#881B1B"), 
                    name = "Turbidity (NTU)" )+
  #theme_cowplot()+
  theme(panel.background = element_rect(fill = "white", colour = "black"))+
  ylim(60, 170)+
  scale_x_continuous(breaks=c(0, 15, 29, 50), limits = c(-1, 57))+
  labs(fill = "Turbidity (NTU)")+
  xlab("Chamber Turbidity (NTU)")+
  ylab("ETRm")+
  ggtitle("ETRm By Turbidity Treatment")

p2
```
There might be another way to look at this
```{r}
#And there's probably an elegant solution with ifelse, but we can also do it with #gsub
fluoro$At.or.Below.Legal.Threshold <- fluoro$Treatment
fluoro$At.or.Below.Legal.Threshold <- as.character(fluoro$At.or.Below.Legal.Threshold) #Make it character

#Now we make our substitutions
fluoro$At.or.Below.Legal.Threshold <- gsub("50", "No", fluoro$At.or.Below.Legal.Threshold) #And we do 50 first to avoid problems with
#zeros
fluoro$At.or.Below.Legal.Threshold <- gsub("0", "Yes", fluoro$At.or.Below.Legal.Threshold)
fluoro$At.or.Below.Legal.Threshold <- gsub("15", "Yes", fluoro$At.or.Below.Legal.Threshold)
fluoro$At.or.Below.Legal.Threshold <- gsub("29", "Yes", fluoro$At.or.Below.Legal.Threshold)

#Now we make it a factor
fluoro$At.or.Below.Legal.Threshold <- 
  as.factor(fluoro$At.or.Below.Legal.Threshold)
```

```{r}
#So now we make our Bayesian Models
bayes_rm.factor <- generalTestBF(ETRm ~ At.or.Below.Legal.Threshold + Genotype, 
                          data = fluoro, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm.factor
```
I don't think this is more likely, but let's plot it to be sure
```{r}
plot(bayes_rm.factor)
```
Alright. Less likely than just genotype.

There might be one last way to do this
```{r}
#And there's probably an elegant solution with ifelse, but we can also do it with #gsub
fluoro$At.or.Above.Legal.Threshold <- fluoro$Treatment
fluoro$At.or.Above.Legal.Threshold <- as.character(fluoro$At.or.Above.Legal.Threshold) #Make it character

#Now we make our substitutions
fluoro$At.or.Above.Legal.Threshold <- gsub("50", "Yes", fluoro$At.or.Above.Legal.Threshold) #And we do 50 first to avoid problems with
#zeros
fluoro$At.or.Above.Legal.Threshold <- gsub("0", "No", fluoro$At.or.Above.Legal.Threshold)
fluoro$At.or.Above.Legal.Threshold <- gsub("15", "No", fluoro$At.or.Above.Legal.Threshold)
fluoro$At.or.Above.Legal.Threshold <- gsub("29", "Yes", fluoro$At.or.Above.Legal.Threshold)

#Now we make it a factor
fluoro$At.or.Above.Legal.Threshold <- 
  as.factor(fluoro$At.or.Above.Legal.Threshold)
```


```{r}
#So now we make our Bayesian Models
bayes_rm.high <- generalTestBF(Fv.Fm ~ At.or.Above.Legal.Threshold + Genotype, 
                          data = fluoro, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm.high
```
I don't think this is more likely, but let's plot it to be sure
```{r}
plot(bayes_rm.high)
```
Hoo boy
```{r}
#Gets every model diveded by evey other model
bfmat <- bayes_rm.high/bayes_rm.high
look.at.this <- bfmat[,2] #Prints out the 2nd column, which should show everything divided by genotype

look.at.this
```
4.64 times as likely. This feels like the Bayesian equivalent of p-hacking, but here we are.

And I guess we try to do something 'clever' with the graph
```{r}
#And I guess we need to bring these in if we want the thing to work
source("R_rainclouds.R") #With the sources we need to copy them into the same project for unfortunate reasons
library(cowplot)
library(readr)
library(ggplot2)
library(dplyr)
library(ggrepel)


#And then we make our graph
p2 <- ggplot(fluoro, aes(x = At.or.Above.Legal.Threshold, y = Fv.Fm, label = Genotype))+
  geom_flat_violin(data= fluoro, aes(x= At.or.Above.Legal.Threshold, 
                                          y = Fv.Fm, 
                                        fill = factor(Treatment)), 
                   position = position_nudge(x = 0, y = 0), adjust = 1.5, trim = TRUE, alpha = .5, colour = NA)+
  geom_text_repel(aes(x= At.or.Above.Legal.Threshold, y = Fv.Fm, 
                                    colour = factor(Treatment)), 
                  nudge_x = -0.3, show.legend = FALSE)+
  scale_color_manual(values = c("black", "#1E90FF", "#000080", "#881B1B"))+
  scale_fill_manual(values = c("black", "#1E90FF", "#000080", "#881B1B"))+
  #theme_cowplot()+
  theme(panel.background = element_rect(fill = "white", colour = "black"))+
  #ylim(10, 190)+
  labs(fill = "Target NTU")+
  xlab("Turbidity Level")+
  scale_x_discrete(labels = c(No = "Less than 29", Yes = "29 or Greater"))+
  ylab("Fv/Fm")+
  ggtitle("Fv/Fm By Turbidity Levels")

p2
```

saves our graph, which is useful for the future
```{r}
ggsave(p2, 
       filename = "Fv/FMPhase1After.png",
       device = "png")
```

