---
title: "YNPQ After Experiment"
author: "Abby Cannon"
date: "2024-11-07"
output: html_document
---

```{r}
#Let's bring in our data
datos <- read.csv("OFAV_PAM_P1_11Jun2024.csv")

#Looks good, but we need to get rid of the sacrifices
class(datos$Treatment) #Character
unique(datos$Treatment) #We need to get rid of "x" 

#Making a new dataframe without the sacrifice treatments
datos.no.sac <- datos[ which(datos$Treatment != "x"), ]

#Now we need to make our days more quantitative
dias <- read.csv("RecoveryPhaseDays.csv")

datos.days <- merge(datos.no.sac, dias, by = "Recovery.Phase")
```

Then we select for only the After data
```{r}
after.datos <- datos.days[ which(datos.days$Days.Post.Experiment == 0), ]
```

We might be able to leave treatment as is, depending on how it's classified
```{r}
class(after.datos$Treatment) #It's character. That's a problem.
Treatment <- as.numeric(after.datos$Treatment) #That's how we fix it
after.datos$Treatment <- Treatment
```

And now that we know that the replicates are of the same coral fragment, we should put them together to avoid pseudoreplication
```{r}
fluoro <- aggregate(Y.NPQ. ~ Coral.ID + Treatment + Genotype, 
                    data = after.datos, mean)
```

Now it's time to start playing with data
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

And we should probably also consider Genotype as a potential factor to avoid pseudoreplication
```{r}
#And it needs to be a factor
fluoro$Genotype <- as.factor(fluoro$Genotype)
```


```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(Y.NPQ. ~ Treatment + Genotype, 
                          data = fluoro, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

```{r}
bayes_rm
```

