---
title: "ETRmBayesianLinearModelAfter"
author: "Abby Cannon"
date: "2024-10-30"
output: html_document
---

First we bring in our data
```{r}
datos <- read.csv("ETRmTreatmentDays.csv")
```

Then we select for only the After data
```{r}
after.datos <- datos[ which(datos$Days.Post.Experiment == 0), ]
```

Then we account for temperature and turbidity and make them numeric
```{r}
after.datos$Temperature <- after.datos$Treatment
class(after.datos$Temperature) #Already character
#So now we can make some substitutions to get it what it needs to be
unique(after.datos$Temperature) #Figuring our what we're actually substituting
#"Control"            "High NTU"           "High Temp"          "High Temp High NTU"

#Now we make the substitutions
after.datos$Temperature <- gsub("High Temp High NTU", "30", after.datos$Temperature)
#And we have to do high temp high NTU first, because it has the names of others in it
after.datos$Temperature <- gsub("Control", "27", after.datos$Temperature)
after.datos$Temperature <- gsub("High NTU", "27", after.datos$Temperature)
after.datos$Temperature <- gsub("High Temp", "30", after.datos$Temperature)

#Now we make it numeric
temperatura <- as.numeric(after.datos$Temperature)

#And then we put it back in our data frame
after.datos$Temperature <- temperatura
```

And now we'll have to do something very similar for turbidity
```{r}
after.datos$Turbidity <- after.datos$Treatment

#Now we make the substitutions
after.datos$Turbidity <- gsub("High Temp High NTU", "29", after.datos$Turbidity)
#And we have to do high temp high NTU first, because it has the names of others in it
after.datos$Turbidity <- gsub("Control", "0", after.datos$Turbidity)
after.datos$Turbidity <- gsub("High NTU", "29", after.datos$Turbidity)
after.datos$Turbidity <- gsub("High Temp", "0", after.datos$Turbidity)

#Now we make it numeric
turbidez <- as.numeric(after.datos$Turbidity)

#And then we put it back in our data frame
after.datos$Turbidity <- turbidez
```

Now it's time to start playing with data
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

We'll also need to make an ID for all of the corals to separate out the random noise
```{r}
#And now for this to work we're going to need an ID column
after.datos$ID <- paste(after.datos$Coral.ID, after.datos$Replicate, sep = "-")
#And it needs to be a factor
after.datos$ID <- as.factor(after.datos$ID)
```

And we should probably also consider Coral.ID as a potential factor to avoid pseudoreplication
```{r}
#And it needs to be a factor
after.datos$Coral.ID <- as.factor(after.datos$Coral.ID)
```
Although I think there aren't enough observations to include it.

There's something fishy going on with our data and I'm trying to figure it out.
```{r}
class(after.datos$ETRm) #It's numeric so that's not the problem
```

Maybe out ID got weird
```{r}
X.factor <- as.factor(after.datos$X)
after.datos$X <- X.factor
```
That is not the problem

Maybe we can try it with the categories as factors and see if that improves anything
```{r}
#after.datos$Temperature <- as.factor(after.datos$Temperature)
#after.datos$Turbidity <- as.factor(after.datos$Turbidity)
```

```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(ETRm ~ Temperature * Turbidity + Coral.ID,
                          data = after.datos, whichRandom = "Coral.ID", neverExclude = "^Coral.ID$")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm
```
```{r}
plot(bayes_rm)
```
So how much more likely than coral ID are we really?
Now we get a matrix of the models to see how much more likely this is than just ID
```{r}
#Gets every model diveded by evey other model
bfmat <- bayes_rm/bayes_rm

look.at.this <- bfmat[,5] #Prints out the 19th column, which should show everything divided by Coral.ID

look.at.this
```
