---
title: "ETRmBayesianLinearModelAfter"
author: "Abby Cannon"
date: "2024-10-30"
output: html_document
---

First we bring in our data
```{r}
datos <- read.csv("ETRmTreatmentDays.csv")
```

Then we select for only the After data
```{r}
after.datos <- datos[ which(datos$Days.Post.Experiment == 0), ]
```

Then we account for temperature and turbidity and make them numeric
```{r}
after.datos$Temperature <- after.datos$Treatment
class(after.datos$Temperature) #Already character
#So now we can make some substitutions to get it what it needs to be
unique(after.datos$Temperature) #Figuring our what we're actually substituting
#"Control"            "High NTU"           "High Temp"          "High Temp High NTU"

#Now we make the substitutions
after.datos$Temperature <- gsub("High Temp High NTU", "30", after.datos$Temperature)
#And we have to do high temp high NTU first, because it has the names of others in it
after.datos$Temperature <- gsub("Control", "27", after.datos$Temperature)
after.datos$Temperature <- gsub("High NTU", "27", after.datos$Temperature)
after.datos$Temperature <- gsub("High Temp", "30", after.datos$Temperature)

#Now we make it numeric
temperatura <- as.numeric(after.datos$Temperature)

#And then we put it back in our data frame
after.datos$Temperature <- temperatura
```

And now we'll have to do something very similar for turbidity
```{r}
after.datos$Turbidity <- after.datos$Treatment

#Now we make the substitutions
after.datos$Turbidity <- gsub("High Temp High NTU", "29", after.datos$Turbidity)
#And we have to do high temp high NTU first, because it has the names of others in it
after.datos$Turbidity <- gsub("Control", "0", after.datos$Turbidity)
after.datos$Turbidity <- gsub("High NTU", "29", after.datos$Turbidity)
after.datos$Turbidity <- gsub("High Temp", "0", after.datos$Turbidity)

#Now we make it numeric
turbidez <- as.numeric(after.datos$Turbidity)

#And then we put it back in our data frame
after.datos$Turbidity <- turbidez
```

And now that we know that the replicates are of the same coral fragment, we should put them together to avoid pseudoreplication
```{r}
fluoro <- aggregate(ETRm ~ Coral.ID + Temperature + Turbidity + Genotype +
                      Treatment, data = after.datos, mean)
```

Now it's time to start playing with data
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

And we should probably also consider Genotype as a potential factor to avoid pseudoreplication
```{r}
#And it needs to be a factor
fluoro$Genotype <- as.factor(fluoro$Genotype)
```


```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(ETRm ~ Temperature * Turbidity + Genotype, 
                          data = fluoro, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm
```

```{r}
plot(bayes_rm)
```

```{r}
#Gets every model diveded by evey other model
bfmat <- bayes_rm/bayes_rm
look.at.this <- bfmat[,5] #Prints out the 19th column, which should show everything divided by ID

look.at.this
```
11 time more likely than just genotype

Now we make our graph to show there's no difference
```{r}
#And I guess we need to bring these in if we want the thing to work
source("R_rainclouds.R") #With the sources we need to copy them into the same project for unfortunate reasons
library(cowplot)
library(readr)
library(ggplot2)
library(dplyr)
library(ggrepel)


#And then we make our graph
p2 <- ggplot(fluoro, aes(x = Temperature, y = ETRm, label = Genotype))+
  geom_flat_violin(data= fluoro, aes(x= Temperature, 
                                          y = ETRm, 
                                        fill = Treatment), 
                   position = position_nudge(x = 0, y = 0), adjust = 1.5, trim = TRUE, alpha = .5, colour = NA)+
  geom_text_repel(aes(x= Temperature, y = ETRm, 
                                    colour = Treatment), 
                  nudge_x = -0.3, show.legend = FALSE)+
  scale_color_manual(values = c("black", "#000080", "#FF0000", "#00CD00"))+
  scale_fill_manual(values = c("black", "#000080", "#FF0000", "#00CD00"))+
  #theme_cowplot()+
  theme(panel.background = element_rect(fill = "white", colour = "black"))+
  #ylim(0.62, 0.725)+
  scale_x_continuous(breaks = c(27, 30), labels = c(27, 30), 
                     limits = c(25.5, 31.5))+
  labs(fill = "Treatment")+
  ylim(50, 170)+
  xlab("Temperature (ºC)")+
  ylab("ETRm (μmol electrons/m^2/s)")+
  ggtitle("ETRm by Treatment")

p2
```

saves our graph, which is useful for the future
```{r}
ggsave(p2, 
       filename = "ETRmPhase2AfterBetterLabelBetterAxesTreatment.png",
       device = "png")
```
