---
title: "ETRm Phase 2 Before"
author: "Abby Cannon"
date: "2024-11-01"
output: html_document
---

First we bring in our data
```{r}
datos <- read.csv("ETRmTreatmentDays.csv")
```

Then we select for only the Before data
```{r}
before.datos <- datos[ which(datos$Days.Post.Experiment == -4), ]
```

Then we account for temperature and turbidity and make them numeric
```{r}
before.datos$Temperature <- before.datos$Treatment
class(before.datos$Temperature) #Already character
#So now we can make some substitutions to get it what it needs to be
unique(before.datos$Temperature) #Figuring our what we're actually substituting
#"Control"            "High NTU"           "High Temp"          "High Temp High NTU"

#Now we make the substitutions
before.datos$Temperature <- gsub("High Temp High NTU", "30", before.datos$Temperature)
#And we have to do high temp high NTU first, because it has the names of others in it
before.datos$Temperature <- gsub("Control", "27", before.datos$Temperature)
before.datos$Temperature <- gsub("High NTU", "27", before.datos$Temperature)
before.datos$Temperature <- gsub("High Temp", "30", before.datos$Temperature)

#Now we make it numeric
temperatura <- as.numeric(before.datos$Temperature)

#And then we put it back in our data frame
before.datos$Temperature <- temperatura
```

And now we'll have to do something very similar for turbidity
```{r}
before.datos$Turbidity <- before.datos$Treatment

#Now we make the substitutions
before.datos$Turbidity <- gsub("High Temp High NTU", "29", before.datos$Turbidity)
#And we have to do high temp high NTU first, because it has the names of others in it
before.datos$Turbidity <- gsub("Control", "0", before.datos$Turbidity)
before.datos$Turbidity <- gsub("High NTU", "29", before.datos$Turbidity)
before.datos$Turbidity <- gsub("High Temp", "0", before.datos$Turbidity)

#Now we make it numeric
turbidez <- as.numeric(before.datos$Turbidity)

#And then we put it back in our data frame
before.datos$Turbidity <- turbidez
```

And now that we know that the replicates are of the same coral fragment, we should put them together to avoid pseudoreplication
```{r}
fluoro <- aggregate(ETRm ~ Coral.ID + Temperature + Turbidity + Genotype, 
                    data = before.datos, mean)
```

Now it's time to start playing with data
```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

And we should probably also consider Genotype as a potential factor to avoid pseudoreplication
```{r}
#And it needs to be a factor
fluoro$Genotype <- as.factor(fluoro$Genotype)
```


```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(ETRm ~ Temperature * Turbidity + Genotype, 
                          data = fluoro, whichRandom = "Genotype", 
                          neverExclude = "Genotype")
```

Prints out our model and tells us what's good.
```{r}
bayes_rm
```
```{r}
plot(bayes_rm)
```
Random effects of genotype are the most likely. Thank goodness.


