---
title: "Nighttime Integrated Threshold Respirometry"
author: "Abby"
date: "2024-08-13"
output: html_document
---

First let's bring in our data files
```{r}
dos <- read.csv("Trial2NoBG.csv")
tres <- read.csv("Trial3NoBG.csv")
cuatro <- read.csv("Trial4NoBG.csv")
nueve <- read.csv("Trial9NoBG.csv")
```

clearning up our data
```{r}
#Getting rid of a column of NAs
dos <- dos[ , -which(names(dos) %in% c("NA."))]
tres <- tres[ , -which(names(tres) %in% c("NA."))]

#Now we need a fuction to get the columns in common
get_common_cols <- function(df1, df2)  intersect(names(df1), names(df2))
compartido <-get_common_cols(dos, tres) #List of the shared ones
library(dplyr) #Package we need for this
tres %>% 
  select(-matches(compartido)) #Drops all the shared ones and only leaves the one that isn't
```

So we have some disagreement about how the day is labeled. Let's rename columns by name
```{r}
names(dos)[names(dos) == 'Day'] <- 'Day.of.Exp'
names(tres)[names(tres) == 'Day'] <- 'Day.of.Exp'
names(cuatro)[names(cuatro) == 'Day'] <- 'Day.of.Exp'
names(nueve)[names(nueve) == 'Day'] <- 'Day.of.Exp'
#Not all of them needed re-naming, but there's no harm in doing it this way and it doesn't take much more programming time
```

Now our problem seems to be that Treatment is classified differently in the different data frames, so we need to fix that.
```{r}
unique(dos$Treatment) #50 29 15  0
unique(tres$Treatment) #50 29 15  0
unique(cuatro$Treatment) #50 29 15  0
unique(nueve$Treatment) #"50 NTU" "29 NTU" "15 NTU" "0 NTU" There's the problem
```

Let's make that one integer
```{r}
numeros <- substr(nueve$Treatment, 1, 2) #Gets the first two pieces of each character
inte <- as.integer(numeros) #Makes them into integers like the rest
nueve$Treatment <- inte
```

Now maybe we can put everything together
```{r}
datos <- bind_rows(dos, tres, cuatro, nueve)
```

How many different values do we have for day and night
```{r}
unique(datos$Time.of.Day) #So we have some trailing white spaces we need gone
```

Getting rid of the white spaces
```{r}
# Returns string without trailing white space
trim.trailing <- function (x) sub("\\s+$", "", x)

#Let's test it
hora <- trim.trailing(datos$Time.of.Day)
unique(hora)
datos$Time.of.Day<- hora
#It worked
```

Now we have to deal with filling in the missing loops for some treatments.
4 trials * 8 chambers * 72 hours = 2304
datos has 2250 rows so we're missing 54 loops total. This may be a bigger or smaller problem depending on where the missing values are.

First we're going to make a dummy data frame for all hours and treatments and chambers and merge it to our actual data to find where our missing values are
```{r}
#sequence for hours
hora.seq <- (1:72)
Hour <- rep(hora.seq, 32) #Repeats our sequence 32 times

#Sequence for chambers
camara.sec <- c(rep(1, 72), rep(2, 72), rep(3, 72), rep(4, 72), rep(5, 72), rep(6, 72),
                rep(7, 72), rep(8, 72))
Chamber <- rep(camara.sec, 4) #Repeats it 4 times

#Sequence for Trials
Trial <- c(rep(2, 576), rep(3, 576), rep(4, 576), rep(9, 576))

ideal <- data.frame(Hour, Chamber, Trial)

datos.with.missing <- merge(datos, ideal, by = c("Hour", "Trial", "Chamber"), all = TRUE)
```

So looking at data there's only one time there was a first hour of the experiment missing. There's also probably a way to interpolate everything, but based on how that command works I may need to restructure the data a bit
```{r}
#Making a less unwieldy data frame to cast
data.for.interp <- data.frame (Trial = datos.with.missing$Trial, Chamber = datos.with.missing$Chamber, Hour = datos.with.missing$Hour, oxy = datos.with.missing$MO2.no.Background.Correction)

#Although it's going to be hard to do with two ID pieces
data.for.interp$ID <- paste("Trial." ,data.for.interp$Trial, ".Chamber.", data.for.interp$Chamber, sep = "")

library(reshape2) #Bringing in the reshape package

#Making a wide data set that only has 72 observations
ancho <- data.for.interp %>%
    dcast(Hour ~ ID, value.var = "oxy")

library(zoo) #We need this package to interpolate
library(dplyr) #And we need this one interpolate across multiple columns at once and fill in some data
library(tidyr)

#-0.0096974659 is the value for hour 2 in 2-7, which is the only one that has an NA as its first value

#So we tell it that it's first value and second value are the same this time
ancho[1,][is.na(ancho[1,])] <- -0.0096974659

#The problem is we have more columns with 72 being blank. But there's a way we can fill them in. This replaces the NAs with the values above them
ancho[71:72,] <- ancho[71:72,] %>% 
  fill(2:33, .direction = "down")
```

Now maybe we can finally interpolate the NAs that aren't on the ends of things
```{r}
ancho.filled <- ancho %>%
  mutate(across(c(Trial.2.Chamber.1:Trial.9.Chamber.8), ~ na.approx(.x, na.rm = TRUE, maxgap=4)))
```

Now we melt our data and put it back together with what we already had
```{r}
melted.ancho <- melt(ancho.filled, id = c("Hour"))

#Let's rename for the sake of our sanity
names(melted.ancho)[names(melted.ancho) == 'value'] <- 'Interpolated.No.BG.MO2'

#And now let's bring back our old columns with some substringing
melted.ancho$Trial <- substr(melted.ancho$variable, 7,7)
melted.ancho$Chamber <- substr(melted.ancho$variable, 17, 17)

#And then some additional renaming that will be useful later
names(melted.ancho)[names(melted.ancho) == 'variable'] <- 'ID'

#Putting it back together
datos.int <- merge(datos.with.missing, melted.ancho, by = c("Hour", "Trial", "Chamber"))
```


Now let's make sure we aren't having overlap between day and night
```{r}
#Checking that we can separate day and night
dia <- datos.int[ which(datos.int$Time.of.Day == "day"), ]
unique(dia$Hour) #Only day values

noche<-datos.int[ which(datos.int$Time.of.Day == "night"), ]
unique(noche$Hour) #Only night values

#Seeing where they overlap
intersect(unique(noche$Hour), unique(dia$Hour)) #They don't. Perfect
```

Now we aggregate by day and for day
```{r}
#So now we have to work out how to calculate areas
library(zoo) #Apparently we need this package
```

```{r}
#And a slightly different function for a slightly different data frame
#Still area under a curve
baja.curva.3 <- function(noche) {
  noche %>%
    summarise(oxy = sum(diff(Hour[order(Hour)])*rollmean(Interpolated.No.BG.MO2[order(Hour)],2)))
     }

#Making a tibble of our data
library(tidyverse)

oxy.by.night <- noche %>%
  group_by(Trial, Treatment, Chamber, Genotype, Day.of.Exp) %>%
  baja.curva.3()

```

Even more packages and then plots and statistical analysis
```{r}
source("R_rainclouds.R") #With the sources we need to copy them into the same project for unfortunate reasons
source("summarySE.R") #With the sources we need to copy them into the same project for unfortunate reasons
library(cowplot)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)


sum.oxy.by.night <- summarySE(oxy.by.night, measurevar = "oxy",
                       groupvars=c("Treatment", "Day.of.Exp"))

p2 <- ggplot(oxy.by.night, aes(x=Day.of.Exp, y = oxy))+
  
  geom_flat_violin(data=oxy.by.night, aes(x=factor(Day.of.Exp, levels = 1:3), y = oxy, 
                                        fill = factor(Treatment)), 
                   position = position_nudge(x = .25, y = 0), adjust = 1.5, trim = TRUE, alpha = .5, colour = NA)+
  geom_point(aes(x=factor(Day.of.Exp, levels = 1:3), y = oxy, 
                                    colour = factor(Treatment), shape = factor(Treatment)), 
             position = position_jitter(width = .2), size = 2)+
  geom_line(data = sum.oxy.by.night, aes(x = as.numeric(Day.of.Exp)+.25, y = oxy_mean, 
                                       group = factor(Treatment), colour = factor(Treatment)), 
            linetype = 3)+
  scale_color_manual(values = c("black", "#1E90FF", "#000080", "#881B1B"))+
  scale_fill_manual(values = c("black", "#1E90FF", "#000080", "#881B1B"))+
  scale_shape_manual(values = c(15, 3, 17, 4))+
  labs(color = "Target NTU", fill= "Target NTU", shape = "Target NTU")+ 
  geom_hline(yintercept = 0, color = "black", linetype = "dashed")+
  #theme_cowplot()+
  theme(panel.background = element_rect(fill = "white", colour = "black"))+
  xlab("Day of Experiment")+
  ylab("Oxygen Produced mg/L/12 hrs")+
  ggtitle("Nighttime Oxygen Production by Day and Treatment")

p2
```
Possibly not much of an effect on anything

But this way we get our image
```{r}
ggsave(p2, 
       filename = "NighttimeThresholdIntegrated.pdf",
       device = "pdf")
```

```{r}
ggsave(p2, 
       filename = "NighttimeThresholdIntegrated.png",
       device = "png")
```

```{r}
# Load packages
library(BayesFactor) # needed to calculate Bayes factors
library(here) # needed to assign working directory for relative paths 
```

```{r}
#And now for this to work we're going to need an ID column
oxy.by.night$ID <- paste(oxy.by.night$Trial, oxy.by.night$Chamber, sep = "-")
#And it needs to be a factor
oxy.by.night$ID <- as.factor(oxy.by.night$ID)

#Now let's see what our treatment is. We want it Numeric or integer
class(oxy.by.night$Treatment) #Integer, probably okay
```

```{r}
#And day of experiment
class(oxy.by.night$Day.of.Exp) #Also integer
```

```{r}
#And genotype
class(oxy.by.night$Genotype) #Integer. We want it to be a factor
```

```{r}
oxy.by.night$Genotype <- as.factor(oxy.by.night$Genotype) #Numbers don't mean anything
```

Now let’s make some models
```{r}
#So now we make our Bayesian Models
bayes_rm <- generalTestBF(oxy ~ Day.of.Exp * Treatment * Genotype + ID,
                          data = oxy.by.night, whichRandom = "ID", neverExclude="^ID$")
```

```{r}
#Prints out our model and tells us what's good.
bayes_rm
```

```{r}
plot(bayes_rm)
```
So Genotype+ID matters more than ID only

#Gets every model diveded by evey other model
```{r}
bfmat <- bayes_rm/bayes_rm

bfmat[,1] #Prints out the first column, which lets us looks for our most likely model. It will have the biggest number
```
Lucky number 5

```{r}
#Gets us a nice data frame to look at for future reference.
c <- as.matrix(bfmat[5,])

x <- as.data.frame(c)
```
10 times more likely than just ID



