---
title: "Daytime Threshold Integrated Offset from Night Max"
author: "Abby Cannon"
date: "2024-10-08"
output: html_document
---

First let's bring in our data files
```{r}
dos <- read.csv("Trial2FromNightMax.csv")
tres <- read.csv("Trial3FromNightMax.csv")
cuatro <- read.csv("Trial4FromNightMax.csv")
nueve <- read.csv("Trial9FromNightMax.csv")
```

clearning up our data
```{r}
#Getting rid of a column of NAs
dos <- dos[ , -which(names(dos) %in% c("NA."))]
tres <- tres[ , -which(names(tres) %in% c("NA."))]

#Now we need a fuction to get the columns in common
get_common_cols <- function(df1, df2)  intersect(names(df1), names(df2))
compartido <-get_common_cols(dos, tres) #List of the shared ones
library(dplyr) #Package we need for this
tres %>% 
  select(-matches(compartido)) #Drops all the shared ones and only leaves the one that isn't
```

So we have some disagreement about how the day is labeled. Let's rename columns by name
```{r}
names(dos)[names(dos) == 'Day'] <- 'Day.of.Exp'
names(tres)[names(tres) == 'Day'] <- 'Day.of.Exp'
names(cuatro)[names(cuatro) == 'Day'] <- 'Day.of.Exp'
names(nueve)[names(nueve) == 'Day'] <- 'Day.of.Exp'
#Not all of them needed re-naming, but there's no harm in doing it this way and it doesn't take much more programming time
```

Now our problem seems to be that Treatment is classified differently in the different data frames, so we need to fix that.
```{r}
unique(dos$Treatment) #50 29 15  0
unique(tres$Treatment) #50 29 15  0
unique(cuatro$Treatment) #50 29 15  0
unique(nueve$Treatment) #"50 NTU" "29 NTU" "15 NTU" "0 NTU" There's the problem
```
Let's make that one integer
```{r}
numeros <- substr(nueve$Treatment, 1, 2) #Gets the first two pieces of each character
inte <- as.integer(numeros) #Makes them into integers like the rest
nueve$Treatment <- inte
```

Now maybe we can put everything together
```{r}
datos <- bind_rows(dos, tres, cuatro, nueve)
```

How many different values do we have for day and night
```{r}
unique(datos$Time.of.Day) #So we only have "day" and "night", which is perfect
```

Now we have to deal with filling in the missing loops for some treatments.
4 trials * 8 chambers * 72 hours = 2304
datos has 2250 rows so we're missing 54 loops total. This may be a bigger or smaller problem depending on where the missing values are.

First we're going to make a dummy data frame for all hours and treatments and chambers and merge it to our actual data to find where our missing values are
```{r}
#sequence for hours
hora.seq <- (1:72)
Hour <- rep(hora.seq, 32) #Repeats our sequence 32 times

#Sequence for chambers
camara.sec <- c(rep(1, 72), rep(2, 72), rep(3, 72), rep(4, 72), rep(5, 72), rep(6, 72),
                rep(7, 72), rep(8, 72))
Chamber <- rep(camara.sec, 4) #Repeats it 4 times

#Sequence for Trials
Trial <- c(rep(2, 576), rep(3, 576), rep(4, 576), rep(9, 576))

ideal <- data.frame(Hour, Chamber, Trial)

datos.with.missing <- merge(datos, ideal, by = c("Hour", "Trial", "Chamber"), all = TRUE)
```

So looking at data there's only one time there was a first hour of the experiment missing. There's also probably a way to interpolate everything, but based on how that command works I may need to restructure the data a bit
```{r}
#Making a less unwieldy data frame to cast
data.for.interp <- data.frame (Trial = datos.with.missing$Trial, Chamber = datos.with.missing$Chamber, Hour = datos.with.missing$Hour, oxy = datos.with.missing$MO2.from.Dark.Max)

#Although it's going to be hard to do with two ID pieces
data.for.interp$ID <- paste("Trial." ,data.for.interp$Trial, ".Chamber.", data.for.interp$Chamber, sep = "")

library(reshape2) #Bringing in the reshape package

#Making a wide data set that only has 72 observations
ancho <- data.for.interp %>%
    dcast(Hour ~ ID, value.var = "oxy")

library(zoo) #We need this package to interpolate
library(dplyr) #And we need this one interpolate across multiple columns at once and fill in some data

#-0.0011533543 is the value for hour 2 in 2-7, which is the only one that has an NA as its first value

#So we tell it that it's first value and second value are the same this time
ancho[1,][is.na(ancho[1,])] <- -0.0011533543

#The problem is we have more columns with 72 being blank. But there's a way we can fill them in. This replaces the NAs with the values above them
ancho[71:72,] <- ancho[71:72,] %>% 
  fill(2:33, .direction = "down")
```

