---
title: "Trial7Chamber1NoBG"
author: "Abby"
date: "2024-08-15"
output: html_document
---

Bringing in our raw data and the packages we need
```{r}
#setwd(" ")
library(knitr)
library(ggplot2)
library(cowplot)
library(respR)
library(tidyr)
raw.data <- read.csv("Trial7Resp.csv") #load raw trial data

#So I don't think we need the background data now, based on the adjustment method we use

# define flush and measure periods
flush <- 120 # 2 min flush
wait <- 120   # 2 mins wait
measure <- 3360  # 56 mins measure
```

## Data QA/QC
This chunk is grabbing the columns we need, and changing the column names so that it is in usable form for respR
```{r}
#First Enter your Experiment and chamber 
Trial <- 7 #Experiment labeled to prevent subsequent confustion
Cham <- 1 #Chamber labeled to prevent subsequent confusion

#Gives us a character based on the chamber we want
nombre <- paste("CH", Cham, ".O2...", sep = "")

#Makes it into a tibble, which lets us paste things more easily
datos.crudos <- as_tibble(raw.data)

## to subset: data[x,y] you need the , to make it complete. to leave your X blank while filling in Y means you want all the rows associated with your chosen collumn, and vice-cerse
coral.resp <- datos.crudos  %>% select(X, starts_with(nombre), Phase, Time, Month, Day, Year) 
#AutoResp names the columns "CH1.02...air.sat.." in the raw data set, you have to process the data one chamber at a time, this isolates the raw data for Chamber 1

##renaming the columns to use with RespR
colnames(coral.resp)[2] = "Oxy"

#plots background and trial data for the first chamber
ggplot(coral.resp, aes(x=X, y=Oxy)) + geom_point()
```
Also appears to show something of a decrease

Gets us only measurements as opposed to flushing and waiting
```{r}
#And now something for later matching
measurement.only <- coral.resp[grep("M", coral.resp$Phase),] #Gets us only the ones that start with M
measurement.only$Phase <- gsub("M", "", measurement.only$Phase) #Gets rid of the Ms
#Gives us only the start values of a particular measurement phase
measurement.first <- measurement.only[match(unique(measurement.only$Phase),
                                           measurement.only$Phase),]
```

Now we have to get into some stuff for easier future programming
```{r}
#Selects for only the first observation

names(measurement.first)[names(measurement.first) == 'Time'] <- 'Start.Time'
#Changes the name of the time column for clarity when we ultimately have to put sheets
#together

names(measurement.first)[names(measurement.first) == 'Phase'] <- 'loop'
#Gives it the same name as the column in subsequent data = to make later matching possible

measurement.first$loop <- as.numeric(measurement.first$loop) #Makes it numeric, which
#we'll need for later management
```
86 loops this time

## Start of RespR

### Looking at the data
This is the preliminary view of the entire run of one coral. The default width is 0.1, but you can change that by adding width=X.XX inside the ()
```{r}
##We want rate.rev to be false so that production is positive, and consumption is negative
coral.resp_inspect <- inspect(coral.resp, rate.rev = FALSE)
#The warning about uneven time can be safely disregarded as it stems from the per minute recording 
#interval being converted to decimalised hours, and we can see the minimum and maximum 
#intervals indicate there are no major gaps. 

# This should be similar to the above "test" just with the bump removed, and the width changed 0.2

coral.resp_inspect2 <- inspect(coral.resp, rate.rev = FALSE, width = 0.2)

#Manually subsetting every replicate would be difficult, but since we know they cycle every hour (3600s) 
#and there are further replicates after the first ends at row XXX, we can use straightforward R syntax to perform a loop.
```
Looks like less photosynthesis and respiration over time

```{r}

#so there may be a way to find this programmatically without poking through the data
#every time
m.start <- as.numeric(rownames(as.data.frame(coral.resp)[which(as.data.frame(coral.resp)$Phase =="F1   ", arr.ind=TRUE)[1],]))-120
#Selects for the row where F1 is and subracts 120 from that number
#But we have to tell it to treat it as a dataframe, not a tibble and I assume there are probably reasons for this

m.end <- as.numeric(rownames(as.data.frame(coral.resp)[which(as.data.frame(coral.resp)$Phase=="M86  ", arr.ind=TRUE)[1],]))
#Selects for the first row of the last measurement. Unfortunately this will vary by
#trial

#Goes up by the time
loops <- seq(m.start, m.end, 3600)

## data starts - apply wait and flush periods
starts <- loops + wait + flush #I don't think these actually get used
## data ends - apply wait and flush AND measure periods
ends <- loops + wait + flush + measure #I don't think these actually get used either

```

```{r}
table(coral.resp$Phase)
```

### The Loop function
First we set up an empty list that will store all the rates. The loop function will go through our start times to the end times and find the linear increase/decrease for each cycle. It will store the rate of change (i.e. slopes) into the rmr list
```{r}

#rates_raw <- calc_rate.int(t10_1, 
                  #     starts=loops,
                   #    wait=240,
                    #   measure=3360,
                     #  quiet = TRUE,
                      # legend=TRUE)


rates <- calc_rate.int(coral.resp, 
                       starts=loops,
                       wait= 240, 
                       measure=3360,
                       quiet = TRUE,
                       legend=TRUE,
                       pos=c(1:80)) #viewing each cycle's oxygen data
```
15 is weird

Now we bring in some data to try to not have to re-program the weights and volumes every
time
```{r}
#First we bring in our data
coral.data <- read.csv("Phase2CoralPhysiology.csv")
c <- as.numeric(coral.data$Chamber) #Then we change our character data to numeric
coral.data$Chamber <- c #And change the column

library(data.table) #The package we need

#Making a data table with Experiment, Chamber, and Wet weight for faster recall
d.wet.weight <-data.table(
   Trial = coral.data$Trial,
    Chamber = coral.data$Chamber,
    Adj.WW = coral.data$Adj.WW
)
setkey(d.wet.weight, Trial, Chamber) #Apparently this is important, but I don't 
#know why

#Making a function to get the wet weight for from the coral data sheet
getWW <-function(c1, c2, data=d.wet.weight) {
    d.wet.weight[.(c1, c2)][, Adj.WW]
}

#Gets it based on the Experiment and Chamber values we put at the start
WW <- getWW(Trial, Cham)

#Now the same thing, but for volume
#Making a data table with Experiment, Chamber, and Volume for faster recall
d.volume <-data.table(
   Trial = coral.data$Trial,
    Chamber = coral.data$Chamber,
    Adj.Vol = coral.data$Adj..Volume
)
setkey(d.volume, Trial, Chamber)

#There might be something fishy with what the volume is being retrieved as
class(d.volume$Adj.Vol)
unique(d.volume$Adj.Vol)

#Making a function to get the volume from the data table
getVol <-function(c1, c2, data=d.volume) {
    d.volume[.(c1, c2)][, Adj.Vol]
}

Vol <- getVol(Trial, Cham) #Gives us the value we're looking for

#Now the same, but for ID, which gets us genotype
d.ID <-data.table(
   Trial = coral.data$Trial,
    Chamber = coral.data$Chamber,
    ID = coral.data$ID
)
setkey(d.ID, Trial, Chamber)

getID <-function(c1, c2, data=d.ID) {
    d.ID[.(c1, c2)][, ID]
}

ID <- getID(Trial, Cham)

#Now for treatment
d.Treatment <-data.table(
   Trial = coral.data$Trial,
    Chamber = coral.data$Chamber,
    Treatment = coral.data$Treatment
)
setkey(d.Treatment, Trial, Chamber)

getTreatment <-function(c1, c2, data=d.Treatment) {
    d.Treatment[.(c1, c2)][, Treatment]
}

Treatment <- getTreatment(Trial, Cham)
```

Now we convert our rates based on the mass of the coral
```{r}
rates_con <- convert_rate(rates, 
                          oxy.unit = "%Air", 
                          time.unit = "sec",
                          output.unit = "mg/h/g",
                          volume = 603-Vol, #total resp volume - coral volume (the volume of water in the entire recirculating system) 
                          mass = WW, #wet weight of the coral/organism (make sure to subtract an average weight of a coral plug)
                          S=35, #salinity
                          t=30, #temperature
                          p=1.009)|> #air pressure in hPa
  select_rate(method = "rsq", n = c(0.0,1)) #selecting for r2 
```

```{r}
summary(rates_con)
View(rates_con)
plot(rates_con, type="rate")
```

Mostly respiring

Making our graph
```{r}
mo2 <- rates_con[["summary"]][['rate.output']]
mo2
loop <- rates_con[["summary"]][['rep']]
data_c1 <- data.frame(loop, mo2)

ggplot(data_c1, aes(x=loop, y=mo2)) + geom_point() + theme_classic() +
  ggtitle("OFAV Trial 7 Chamber 1") + geom_hline(aes(yintercept=0)) +
  labs(x="Loops", y=expression(MO[2]~(mgO[2]~kg^-1~h^-1))) 
```
Mostly respiration and some cycles

Now let's clean up so we can combine. That's where Briana's previous used and not used loops come in
```{r}
#Bringing in the old data
old.results <- read.csv("Trial7Results.csv")

#There may be a cheeky way to rename our data and merge it
colnames(data_c1) = c("Cycle", "MO2 no Background Correction")

#But the old results have multiple chambers, so we'll need to add that to our background data
data_c1$Chamber <- rep(Cham)

#Now we put them together, which should keep the information we need
corrected.data <- merge(data_c1, old.results, by = c("Cycle", "Chamber"))

corrected.data$Trial <- rep(Trial) #We'll need this when we put it all together

#Write a csv of our data. We'll put them all together later
#write.csv(corrected.data, "Trial7Chamber1NoBG.csv")
#After we adjust the rates by more correct value
```

This continues to be confusing
```{r}
ggplot(corrected.data, aes(MO2, `MO2 no Background Correction`)) + geom_point() + geom_smooth(method = "lm")
```

So if something is supposed to look like this, something got pretty weird in trial 8