---
title: "Trial 2 Adjusted Background"
author: "Abby"
date: "2024-09-03"
output: html_document
---

So first we bring in our data
```{r}
background <- read.csv("BackgroundRatesCalculated.csv")
MO2.no.BG <- read.csv("Trial2NoBG.csv")
colnames(background) #So there's a lot, but I can make it work
```

So I think it's going to need to start cleaning, because this is a mess.
```{r}
#First lets get our treatment effects
XX <- background[ , grepl( "X" , names( background ) ) ]

#And then we need our loops
Loops <- background$Loop

#Now we put it together
library(dplyr) #bringing in the package we need
LoopXX <- bind_cols(XX, Loops) #new combined data frame

#But it ended up with a truly stupid column name
colnames(LoopXX)[8] <- "Loop"

#And I think a column labeled X is just going to get in the way
LoopXX <- LoopXX[,!names(LoopXX) %in% c("X")]
```

Alright. Now I think we may need to add some information to the background
```{r}
#First we make a column for temperature
MO2.no.BG$Temp <- rep(27)

#Now I think we need to get rid of certain columns with non-relevant info for the sake of my sanity
MO2.no.BG <- MO2.no.BG[,!names(MO2.no.BG) %in% c("X.1", "X", "NA.")]
```

Now we need to work on separating the treatments and temperatures out of our Loops and treatment backgrounds
```{r}
library(reshape2) #bringing in the package we need to melt
meltyloops <- melt(LoopXX, id = "Loop") #Makes it so we have three columns, with loop and our treatments, which are all called variable at the moment and Value is our MO2, which we need to fix

#Renames the column
colnames(meltyloops)[colnames(meltyloops) == 'value'] <- 'Background.MO2'

#Now we need our treatments and temperatures
class(meltyloops$variable) #It's a factor, we many need to change it to character
v <- as.character(meltyloops$variable) #Looks good
meltyloops$variable <- v

#Now we work on extracting our information, which is why we bothered to make it character
library(stringr)
Treatment <- str_match(meltyloops$variable, "X\\s*(.*?)\\s*N") #So that kind of worked, but gave us some extra characters. We can fix it with gsub though
Treatment <- gsub("X", "", Treatment) #First we get rid of the "X"
Treatment <- gsub("N", "", Treatment) #Then we get rid of the "N"
meltyloops$Treatment <- Treatment[,1] #It had 2 for some reason, now we have 1

#Now let's get the second to last characters
Temp <- str_sub(meltyloops$variable, -3, -2) #numbering still goes left to right
meltyloops$Temp <- Temp

```

Now we need to see if any of the background readings trend any particular way
```{r}
library(ggplot2) #Bringing in the package we need for plotting

# Scatter plot by group
ggplot(meltyloops, aes(x = Loop, y = Background.MO2, color = variable)) +
  geom_point()+
  geom_smooth(method = lm)+
  theme_bw() #Easier to make sense of confidence intervals with black and white theme
```
So we'll tentatively say that the 30C treatments decline over time, but the 27C treatments hold steady. So we can take an average of the 27C treatments, but will need to calculate slopes for 30C

```{r}
#First we need to get just the data for the 30C treatments and their respective turbidities

#Just 30 and 0
X0NTU.30C <- meltyloops[ which(meltyloops$variable== "X0NTU.30C"), ]

#Just 30 and 29
X29NTU.30C <- meltyloops[ which(meltyloops$variable== "X29NTU.30C"), ]

#Now we do our linear regressions. 
#We'll start with 30C, 0 NTU
lm.hot.clean <- lm(X0NTU.30C$Background.MO2 ~ X0NTU.30C$Loop)
summary(lm.hot.clean) #MO2 = 0.01188*Loop -0.2015851
#Now 30C, 29NTU
lm.hot.dirty <- lm(X29NTU.30C$Background.MO2 ~ X29NTU.30C$Loop)
summary(lm.hot.dirty) #MO2 = 0.005439*Loop -0.074406
```
So this actually gets easier if we don't use the 30C treatments and since this data set is only the threshold experiment we can do that. Although I guess we can get means for everything
```{r}
#Gets us a data frame of all the 
background.means <-aggregate(meltyloops$Background.MO2, list(meltyloops$Treatment, meltyloops$Temp), FUN = "mean")

#Although it looks like we'll need to change some names
colnames(background.means) #"Group.1" "Group.2" "x"

library(data.table) #A package we'll need to easily rename columns

#Changing the names
setnames(background.means, old = c("Group.1", "Group.2", "x"), 
         new = c("Treatment", "Temp", "Background MO2"))

#Now we put it back together and calculate
MO2.and.BG <- merge(MO2.no.BG, background.means, by = c("Treatment", "Temp"))
```

Let's see if this solves the apparent nighttime oxygen production problem or not
```{r}
#Let's subtract chamber oxygen production from coral oxygen production
MO2.and.BG$Coral.MO2 <- MO2.and.BG$MO2 - MO2.and.BG$`Background MO2`
#So that doesn't look good at all, but I'll graph it to prove a point.
```


